% intro.tex

% intro to problems of learning and classification
There has been long-standing interest in constructing systems capable of
learning from experience since even before computers in their modern form have
existed. In the last few decades, with computing power skyrocketing
exponentially coupled with leaping advances in theory of learning systems and
statistical inference, these problems became tractable and eventually came into
use ubiquitously. With applications ranging from facial detection systems for
surveillance to identifying cosmic objects for cosmology, they have found
widespread adoption in industry and academia. These systems circumvent the need
to produce a precise mathematical model for the problem at hand, exploiting
general techniques to instead infer a model from available data. With their
advent, however, has come an ever rising need for computing power to facilitate
their operation. This has found data centers of unprecedented scales consuming
enormous amounts of power to provide the instant predictions we've come to rely
on.

% intro to how quantum may solve it
With snowballing energy and space requirements of classical computers in the
form of GPU clusters and Application Specific Integrated-Circuits (ASICs), there
has been a spark of interest in offloading this computation onto quantum
computers, which, till recently, have largely remained a rare species spotted
only in labs surrounded by helium-cooled superconductors and white-coated
predators. Current scales of available quantum computers, however, still lack
the power required to fully tackle these challenges while maintaining reliable
error-levels or adding their own error checking and correction. This has
motivated using quantum computers to run bottle necked computational subroutines
with classical control systems. These systems generally lack error correction,
and thus earn themselves the title of `noisy'. These form the basis of
computation considered in this thesis, Noisy Intermediate-Scale Quantum (NISQ)
computers.

These methods have been applied to a myriad of problems where a quantum
advantage is expected, and have earned themselves the name `Variational Quantum
Algorithms (VQAs)'. VQAs are characterized by parametrized quantum circuits,
wherein the parameters are controlled by a classical computer running an
optimization routine, updating them based on the measurement outcomes of the
circuit. The technique has shown great promise and bypasses several of the
issues arising from the lack of capability for error checking and correction, in
turn arising from the small number of available qubits, in quantum computers
expected to be manufactured within the next few years. The bulk of the
computation requiring memory, such as parameter updation and gradient
computation for optimization is generally performed on the classical controller.

However, connecting a quantum computer to a classical puppeteer is not expected
to come without its own issues either. It constrains the architecture and is
itself bottle necked on both ends, first by the parameter transfer and
configuration from the classical to the quantum, then finally by the detectors
on the quantum side to the classical. In this thesis, we focus on the former,
discussing the limits of computation and computational precision achievable with
this hybrid architecture.

% report structure
\subsection{Structure}
In \autoref{sec:prelim}, definitions and relevant results in classical
computing, physics, and quantum information are presented.  \autoref{sec:vqa}
reviews Variational Quantum Algorithms and their architecture, while
\autoref{sec:infolimits} discusses bounds on their computational precision
imposed by the architecture. Finally, \autoref{sec:qsvm} discusses Quantum
Support Vector Machines and their error probabilities.

\subsection{Outline of New Results}
We present the classical classification problem, the quantum SVM, and discuss
how error bounds on parametrized quantum circuits extend to bounds on the error
probabilities of the QSVM. Further work is initiated regarding estimating the
error bounds for PQCs.

